---
title: "ESM263 HW4 results data check"
author: "Casey O'Hara"
date: "1/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(DT)

### directory where assignments are saved
dir_asst <- '~/Desktop/esm263_asst4' 
### If you want to try this on your own .dbf file,
### change this to point to the folder where your .dbf is, then knit it
```

# Figure out which scores change b/c viewsheds

```{r}
view_key <- foreign::read.dbf('out_checks/view_scores_out.dbf') %>%
  janitor::clean_names() %>%
  select(calwnum, view_score = mean)
view01_key <- foreign::read.dbf('out_checks/view_scores_out_view01.dbf') %>%
  janitor::clean_names() %>%
  select(calwnum, view01_score = mean)

view_checks <- view_key %>%
  full_join(view01_key, by = 'calwnum') %>%
  mutate(viewdiff = view_score != view01_score)

```

# get student data

Use `list.files` to identify all the .dbf files, including the full path.  Use `lapply` to quickly read in each one into a list.  We can use the `foreign::read.dbf()` function to read this format, instead of `readr::read_csv()`.

Once we have a list of data frames, then use `bind_rows` to combine all the list elements into one big data frame.

## get data, and check for proper data formatting

```{r}
dbfs <- list.files(dir_asst, pattern = '.dbf$',
                   full.names = TRUE, recursive = TRUE)

dbf_df <- lapply(dbfs, FUN = function(f) {
  df <- foreign::read.dbf(f, as.is = TRUE) %>%
    janitor::clean_names() %>%
    mutate(student = basename(dirname(f)) %>% str_replace('_[0-9].*', ''))
  name_typos <- 'scpre|scores'
  names(df) <- names(df) %>% str_replace(name_typos, 'score')
  
  ### make sure columns are correct
  cols_good <- ncol(df) == 3
  colnames_good <- all(names(df) %in% c('score', 'calwnum', 'file'))
  df <- df %>%
    mutate(cols_good = cols_good,
           names_good = colnames_good)
  if('mean' %in% names(df)) df <- df %>% mutate(score = mean)
  return(df)
}) %>%
  bind_rows()

grades_data_format <- dbf_df %>%
  group_by(student) %>%
  summarize(n_watersheds = sum(!is.na(score)),
            cols_good = first(cols_good),
            names_good = first(names_good))

```

## check against a key

``` {r}
mdl_key <- foreign::read.dbf('keys/mdl_out.dbf') %>%
  janitor::clean_names() %>%
  select(calwnum, key_score = mean) %>%
  full_join(view_checks %>% select(calwnum, viewdiff), by = 'calwnum')

scores_df <- dbf_df %>%
  select(student, calwnum, score) %>%
  full_join(mdl_key, by = 'calwnum') %>%
  mutate(score_match = score == key_score) %>%
  group_by(student) %>%
  summarize(n_scores = sum(!is.na(score)),
         score_match_all = sum(score_match, na.rm = TRUE),
         n_scores_viewdiff = sum(!is.na(score) & !viewdiff),
         score_match_viewdiff = sum(score_match & !viewdiff, na.rm = TRUE),
         pct_all = (score_match_all / n_scores) %>% round(2),
         pct_viewdiff = (score_match_viewdiff / n_scores_viewdiff) %>% round(2))

### set some threshold for correct scores to be "close enough"
to_check <- scores_df %>%
  mutate(checkit = pct_all < 0.8 & pct_viewdiff < 0.8) %>%
  select(student, n_scores, pct_all, pct_viewdiff, checkit)

DT::datatable(to_check)
```

# output checks

These were all used to develop a key (and check against a few student samples) for students to check their own work.

## Viewshed zone sums

```{r, eval = FALSE}
view_dbfs <- list.files('out_checks', pattern = 'view.+.dbf$',
                   full.names = TRUE, recursive = TRUE)
view_key <- foreign::read.dbf('keys/view_sums_out.dbf', as.is = TRUE) %>%
  mutate(method = 'full vals') %>%
  bind_rows(foreign::read.dbf('keys/view_sums_0_1_out.dbf', as.is = TRUE) %>%
              mutate(method = 'vals 0 1')) %>%
  janitor::clean_names() %>%
  select(calwnum, key_score = mean, method)

view_df <- lapply(view_dbfs, FUN = function(f) {
  df <- foreign::read.dbf(f, as.is = TRUE) %>%
    janitor::clean_names() %>%
    mutate(file = basename(f))
  if('mean' %in% names(df)) df <- df %>% rename(score = mean)
  return(df)
}) %>%
  bind_rows()


view_scores_df <- view_df %>%
  mutate(score = round(score, 3)) %>%
  select(calwnum, score, file) %>%
  group_by(calwnum, score) %>%
  mutate(n_at_this_score = n()) %>%
  ungroup()

view_tol <- 1
view_check_df <- view_scores_df %>%
  left_join(view_key, by = 'calwnum') %>%
  mutate(correct = (abs(score - key_score) <= view_tol)) %>%
  group_by(file, method) %>%
  summarize(n_correct = sum(correct),
            pct_correct = 100 * n_correct / n())


DT::datatable(view_check_df)
```

## Riparian zone sums

```{r, eval = FALSE}
rip_dbfs <- list.files('out_checks', pattern = 'rip.+.dbf$',
                   full.names = TRUE, recursive = TRUE)
rip_key <- foreign::read.dbf('keys/rip_sums_out.dbf', as.is = TRUE) %>%
  janitor::clean_names() %>%
  select(calwnum, key_score = mean)
rip_df <- lapply(rip_dbfs, FUN = function(f) {
  df <- foreign::read.dbf(f, as.is = TRUE) %>%
    janitor::clean_names() %>%
    mutate(file = basename(f))
  if('mean' %in% names(df)) df <- df %>% rename(score = mean)
  return(df)
}) %>%
  bind_rows()


rip_scores_df <- rip_df %>%
  mutate(score = round(score, 3)) %>%
  select(calwnum, score, file) %>%
  group_by(calwnum, score) %>%
  mutate(n_at_this_score = n()) %>%
  ungroup()

rip_tol <- 1
rip_check_df <- rip_scores_df %>%
  full_join(rip_key, by = 'calwnum') %>%
  mutate(correct = (abs(score - key_score) < rip_tol)) %>%
  group_by(file) %>%
  summarize(n_correct = sum(correct),
            pct_correct = 100 * n_correct / n())


DT::datatable(rip_check_df)
```


## Developable zone sums

```{r, eval = FALSE}
dev_dbfs <- list.files('out_checks', pattern = 'dev.+.dbf$',
                   full.names = TRUE, recursive = TRUE)
dev_key <- foreign::read.dbf('keys/dev_scores_out.dbf', as.is = TRUE) %>%
  janitor::clean_names() %>%
  select(calwnum, key_score = mean)
dev_df <- lapply(dev_dbfs, FUN = function(f) {
  df <- foreign::read.dbf(f, as.is = TRUE) %>%
    janitor::clean_names() %>%
    mutate(file = basename(f))
  if('mean' %in% names(df)) df <- df %>% rename(score = mean)
  return(df)
}) %>%
  bind_rows()


dev_scores_df <- dev_df %>%
  mutate(score = round(score, 3)) %>%
  select(calwnum, score, file) %>%
  group_by(calwnum, score) %>%
  mutate(n_at_this_score = n()) %>%
  ungroup()

dev_tol <- 0
dev_check_df <- dev_scores_df %>%
  full_join(dev_key, by = 'calwnum') %>%
  mutate(diff = abs(score - key_score),
         diff_pct = diff / key_score,
         correct = (diff <= dev_tol)) %>%
  group_by(file) %>%
  summarize(n_correct = sum(correct),
            pct_correct = 100 * n_correct / n())


DT::datatable(dev_check_df)
```

# Make a check key

Join the key tables.  Select 10 or so watersheds and provide scores for rasters along the way as a test.

```{r, eval = FALSE}
dev_key <- foreign::read.dbf('keys/dev_scores_out.dbf', as.is = TRUE) %>%
  janitor::clean_names() %>%
  select(calwnum, develop_score = mean)
rip_key <- foreign::read.dbf('keys/rip_sums_out.dbf', as.is = TRUE) %>%
  janitor::clean_names() %>%
  select(calwnum, riparian_sum = mean)
view_key <- foreign::read.dbf('keys/view_sums_out.dbf', as.is = TRUE) %>%
  janitor::clean_names() %>%
  select(calwnum, view_sum = mean)
# view_key_1_0 <- foreign::read.dbf('keys/view_sums_0_1_out.dbf', as.is = TRUE) %>%
#   select(method = 'vals 0 1')) %>%

set.seed(11235813)
key_df <- dev_key %>%
  full_join(rip_key, by = 'calwnum') %>%
  full_join(view_key, by = 'calwnum') %>%
  filter(!is.na(view_sum) & !is.na(develop_score)) %>%
  sample_n(12) %>%
  arrange(calwnum)
write_csv(key_df, 'keys/key_10_watersheds.csv')

```

