---
title: "HW3 raster grader"
author: "Casey O'Hara"
date: "2/24/2020"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(raster) 
library(tidyverse)
### load raster before tidyverse! otherwise the raster::select (which I
### never use) masks dplyr::select, which I use all the time
library(DT)

### directory where assignments are saved
# dir_asst <- '~/Desktop/esm263_asst3' 
dir_asst <- 'H:/ESM263/HW3_submissions'

### directory where .gdbs are stored
dir_gdb  <- '~/Desktop/esm263_asst3_gdb'

```

# Unzip the .7z files to get the .gdb

```{r, eval = FALSE}

zips <- list.files('H:/ESM263/HW3_submissions', pattern = '.7z$',
                    full.names = TRUE, recursive = TRUE)

for(z in zips) {
  # z <- zips[1]
  gdb_file <- list.files(dirname(z), pattern = '.gdb', full.names = TRUE)
  if(length(gdb_file) > 0) {
    # message('File exists: ', gdb_file, '\n... deleting existing to overwrite!')
    # unlink(gdb_file, recursive = TRUE, force = TRUE)
    message('File exists: ', gdb_file, '\n... skipping extract!')
    unlink(gdb_file, recursive = TRUE, force = TRUE)
    # next()
  }
  message('Unzipping ', basename(z))
  system2(command = 'C:\\Program Files\\7-Zip\\7z.exe',
          args = sprintf('x "%s" -o"%s/"', z, dirname(z)))
}
```

# Get rasters out of geodatabases

Get a list of .gdb paths.  Loop over them.  For each, identify the raster inside, pull it out, save it as a .tif outside the .gdb (in the directory just upstairs from it)

```{r, eval = FALSE}
library(arcgisbinding)
arc.check_product()
dir_asst <- 'H:/ESM263/HW3_submissions'

student_dirs <- list.files(dir_asst)
gdbs <- list.files(dir_asst, pattern = '.gdb$',
                   include.dirs = TRUE,
                   full.names = TRUE, recursive = TRUE)

missing <- student_dirs[!student_dirs %in% basename(dirname(gdbs))]

problems <- paste(c('Janelle', # nothing in gdb
                    'Nicole'),  # nothing in gdb
                  collapse = '|')

gdbs_fixed <- gdbs[!str_detect(gdbs, problems)]

for(gdb in gdbs_fixed) {
  ### gdb <- gdbs[48]
  cat('getting raster from ', gdb, '\n')
  tmp <- arc.open(gdb)
  raster_names <- tmp@children[['RasterDataset']]
  
  if(length(raster_names) > 1) cat('more than one raster!!!!!  ', gdb, '\n  ', paste(raster_names, collapse = ', '), '\n')
  if(length(raster_names) == 0) stop('no raster!!!!!  ', gdb)
  out_file <- file.path(dirname(gdb), paste0(raster_names[1], '.tif'))
  if(file.exists(out_file)) {
    cat('File exists: ', out_file, '... skipping!\n')
    next()
  }
  
  r <- arc.open(file.path(gdb, raster_names[1])) %>%
    arc.raster() %>%
    as.raster()
  
  # plot(r)
  cat('writing to ', out_file, '\n')
  writeRaster(r, out_file, overwrite = TRUE)
}
```

# Check rasters

Use `list.files` to identify all the .tif files, including the full path.  

We can compare each student raster to a key raster (mine... haha).  There should be near perfect agreement, though some variation can be expected if only because some students may have chosen different conversions for miles to meters.  Agreement should look like:

* nearly the same number of suitable cells
* nearly no difference in which cells are noted as suitable

``` {r}
source('helper_fxns.R') ### raster checking functions
```



```{r}
# dirs <- list.files(dir_asst, full.names = TRUE)
tifs <- list.files(dir_asst, pattern = '.tif$',
                   recursive = TRUE, full.names = TRUE)

key_rast  <- raster('hw3_arc_outputs/suitable.tif')

score_df <- data.frame()

for(tif in tifs) {
  # tif <- tifs[46]
  student <- basename(dirname(tif)) %>%
    str_replace('_.*', '')
  
  x <- raster(tif)
  # x; key_rast
  # x <- x %>% projectRaster(key_rast)
  # plot_diffs(x, key_rast)
  values(x)[values(x) <= 0] <- NA
  
  check_df <- check_rast(x, key_rast) %>%
    mutate(student = student)
  score_df <- score_df %>%
    bind_rows(check_df)
}

DT::datatable(score_df)

```
