---
title: "HW3 raster grader"
author: "Casey O'Hara"
date: "2/24/2020"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(raster) 
library(tidyverse)
### load raster before tidyverse! otherwise the raster::select (which I
### never use) masks dplyr::select, which I use all the time
library(DT)

### directory where assignments are saved
dir_asst <- '~/Desktop/esm263_asst3' 

### directory where .gdbs are stored
dir_gdb  <- '~Desktop/esm263_asst3_gdb'

```

# Unzip the .7z files to get the .gdb

```{r, eval = FALSE}
zips <- list.files(dir_asst, pattern = '.7z$',
                    full.names = TRUE, recursive = TRUE)
for(z in zips) {
  # z <- zips[1]
  gdb_file <- list.files(dirname(z), pattern = '.gdb', full.names = TRUE)
  if(length(gdb_file) > 0) {
    message('File exists: ', gdb_file, '\n... skipping extract!')
    # unlink(gdb_file, recursive = TRUE, force = TRUE)
    next()
  }
  message('Unzipping ', basename(z))
  system2(command = '7z',
          args = sprintf('x "%s" -o"%s/"', z, dirname(z)))
}
```


# Check rasters

Use `list.files` to identify all the .img files, including the full path.  

We can compare each raster to one of two rasters: the "suitable" raster and the "suitable regions" raster.  They should actually both just be covering the same cells so shouldn't matter too much.

Unfortunately, only two students have readily identifiable rasters in their Results.gdb, so unless there is another file that is identifiable as a raster format, this is a waste of time... :(

```{r}
# dirs <- list.files(dir_asst, full.names = TRUE)
imgs <- list.files(dir_asst, pattern = '.img$',
                   recursive = TRUE, full.names = TRUE)

good_rast <- x ### need a known "good" raster
for(img in imgs) {
  ### img <- imgs[2]
  x <- raster(img)
  values(x)[values(x) <= 0] <- NA
  # plot(x)
  # check values against good raster, allowing for small differences here and there
}


```