---
title: "HW3 raster grader"
author: "Casey O'Hara"
date: "2/24/2020"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(raster) 
library(tidyverse)
### load raster before tidyverse! otherwise the raster::select (which I
### never use) masks dplyr::select, which I use all the time
library(DT)

### directory where assignments are saved
dir_asst <- '~/Desktop/esm263_asst3' 

### directory where .gdbs are stored
dir_gdb  <- '~Desktop/esm263_asst3_gdb'

```

# Unzip the .7z files to get the .gdb

```{r, eval = FALSE}
zips <- list.files(dir_asst, pattern = '.7z$',
                    full.names = TRUE, recursive = TRUE)
for(z in zips) {
  # z <- zips[1]
  gdb_file <- list.files(dirname(z), pattern = '.gdb', full.names = TRUE)
  if(length(gdb_file) > 0) {
    message('File exists: ', gdb_file, '\n... skipping extract!')
    # unlink(gdb_file, recursive = TRUE, force = TRUE)
    next()
  }
  message('Unzipping ', basename(z))
  system2(command = '7z',
          args = sprintf('x "%s" -o"%s/"', z, dirname(z)))
}
```

# Get rasters out of geodatabases

Get a list of .gdb paths.  Loop over them.  For each, identify the raster inside, pull it out, save it as a .tif outside the .gdb.

Test with my own gdb:
```{r, eval = FALSE}
library(arcgisbinding)
arc.check_product()
dir_asst <- 'H:/ESM263/asst3_submitted'

test_gdb <- 'H:/ESM263/HW3/output.gdb'
tmp <- arc.open(test_gdb)
raster_names <- tmp@children[['RasterDataset']]

if(length(raster_names) > 1) stop('more than one raster!!!!!  ', paste(raster_names, collapse = ', '))
out_file <- file.path(dirname(test_gdb), paste0(raster_names[1], '.tif'))

r <- arc.open(file.path(test_gdb, raster_names[1])) %>%
  arc.raster() %>%
  as.raster()

plot(r)
message('writing to ', out_file)
writeRaster(r, out_file, overwrite = TRUE)
```

# Check rasters

Use `list.files` to identify all the .tif files, including the full path.  

We can compare each student raster to a key raster (mine... haha).  There should be near perfect agreement, though some variation can be expected if only because some students may have chosen different conversions for miles to meters.  Agreement should look like:

* nearly the same number of suitable cells
* nearly no difference in which cells are noted as suitable

```{r}
check_rast <- function(student_rast, key_rast, tol = .002) {
  n_cells_key <- sum(!is.na(values(key_rast)))
  
  cellcount_check <- sum(!is.na(values(student_rast)) - n_cells_key) / n_cells_key %>%
    abs()
  
  ### presumably, extents, CRS, and res should be identical, so stack should
  ### work just fine:
  s_rast_norm <- student_rast / student_rast
  k_rast_norm <- key_rast / key_rast
  tmp_stack <- stack(s_rast_norm, -k_rast_norm)
  
  ### take absolute value of difference 
  diff_rast <- calc(tmp_stack, fun = sum, na.rm = TRUE) %>%
    abs()
  
  celldiff_check <- sum(values(diff_rast), na.rm = TRUE) / n_cells_key
  
  check <- celldiff_check <= tol & cellcount_check <= tol
  
  return(data.frame(good      = check, 
                    celldiff  = celldiff_check, 
                    cellcount = cellcount_check))
}

```


```{r}
# dirs <- list.files(dir_asst, full.names = TRUE)
imgs <- list.files(dir_asst, pattern = '.img$',
                   recursive = TRUE, full.names = TRUE)

key_rast  <- raster('hw3_arc_outputs/suitable.tif')

score_df <- data.frame()

for(img in imgs) {
  ### img <- imgs[2]
  student <- basename(dirname(dirname(dirname(img)))) %>%
    str_replace('_.*', '')
  
  x <- raster(img)
  values(x)[values(x) <= 0] <- NA
  
  check_df <- check_rast(x, key_rast) %>%
    mutate(student = student)
  score_df <- score_df %>%
    bind_rows(check_df)
}

```
